Regression Cube Analysis of Cohort Study Data

### 2015-03-30 draft.tex
resolved TODOs

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
%% TODO: Blind Review: Replace with online ID
\onlineid{175}

@@

%% Paper title.
%% TODO: Replace towards with regarding?

@@

%% TODO: Check DLPD for correct page numbers
While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
\begin{enumerate}
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
% TODO: Do we need a preprocessing section?
%\com{Add Data Preprocessing Section.}
%%
The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
Overview visualizations of feature relationships as presented by Angelelli et al. \cite{Angelelli} are often focused on relationships between the visualized features.
%%
Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
These target features often indicate the presence of the investigated disease.
%%
As described in Section \ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing these relationships.
%%
% unified -> singular [Steffi?]
A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
\subsection{Cube Description Using Regression Formula Notation}
%%
Expert knowledge is integrated into regression analysis by the regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%
%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
%%

@@

\subsection{Abstracting Regression Results Using $R^2$}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.5in]{figures/cube}
 \caption{
 (left) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The $R^2$ values extracted from the regression formulas depict the quality-of-fit and are mapped to color saturation (a saturated color shows a strong correlation).
 %%
 (right) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The $R^2$ value of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%% TODO: Cite another paper here!
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ value is the metric allowing for this kind of assessment.
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between [0,1], it allows for comparison between regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%

@@

\section{System Design} \label{sec:SystemDesign}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.5in]{figures/system}
 \caption{
 Breast fat data set loaded into the \emph{Regression Cube} prototype.
 (a) Using the formula input, the user specifies the dependent feature as well as its calculation rules.
 (b) 3D cube visualization, showing values above the cube matrix diagonal as overview.
 The values of the currently selected slice are mirrored and represented as orange data points on the slicing plane.
 (c) 2D mosaic plot visualization of the selected slice for feature \emph{Pain/Discomfort}.
 %\com{TODO: Adapt formula to final image.}
 %Context information is shown for the regression model of $Parenchym \sim Bronchitis + AlcoholLast30Days$, depicting details about the confidence intervals, coefficients and associated p-values.
 %\com{Screenshot of the cube visualization embedded in the framework}
 %\com{Also include active tooltip}
 }
  \label{fig:System}
\end{figure}
%% scientific english: kurz und langweilig ;)
We designed our system to be freely accessible and easy to use.
With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses, maybe with other (statistical) tools.
Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.
\\\\
% \subsection{Design and Visualization Techniques}
\noindent Design choices and spaces are restricted by the underlying technologies and devices.
%% TODO: Umschreiben
Resting collaboration and exchange about data and methods on web-based technologies shows promising results, as there is no set-up time involved and domain experts can use the methods from any computer connected with the web.
By design, web technology is based on a client-server architecture, making it easy to outsource computationally heavy tasks on server clusters and transferring results to the client device.
The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.
Therefore, we have to adapt our design respecting the affordances of web pages.
\subsection{System Paradigm and Components}
The \emph{Regression Cube} design focuses on a clean interface, reducing the amount of user-interface elements as much as possible.
This allows for a fast learning of the individual system parts.
%More importantly, it allows the user to focus all mental resources on the analysis tasks, rather than wrangling with configuring the system to serve the current task or hypothesis.
Our prototype consists of three components:
	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube [2.SI].
	\item The \emph{formula editor} allows to adjust the formula w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
	It also allows to select a reference formula for creating a difference cube [3.ZF].
%%\com{Focus on simplification and ease of use, reducing the number of UI elements as much as possible (Statistical Visualizations often tend to put all information and settings at once to the user, which is overwhelming for the user).}
\paragraph{File Upload and Classification [1.AF].}
Popular analytics tools, such as WEKA \cite{WEKA}, owe their popularity to their support of open file types.
To allow other users even outside the epidemiological application domain to access to our tool, we use standard ASCII-based CSV files.
The first line in a CSV file represents all features (columns) of the data set.
Each line after that represents one subject (row) and its feature manifestations.

@@

\textbf{Data Security.}
%%
Security issues are raised by uploading data into an online service such as our prototype.
%%
The use of epidemiological data is preceded by a detailed description of the analysis purpose and has to be approved by ethics committees.
%%
Preventive steps have to be taken to restrict access to unauthorized subjects.
%%
%We decided against a user account system because it reduces the ease of use largely without having a direct advantage for the user.
%%
We apply a SHA-256 hashing on the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
%% TODO: Implement this
Data sets are deleted from the server after closing a session.

@@

\paragraph{3D Prism as Data Mini-Map.}
%%
The 3D cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
%%
It serves as a function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
%%
%It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
%%
The displayed prism shows values above the matrix diagonal.
%%
For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes absolute $R^2$ values (Fig.~\ref{fig:System} b).
%%
Applying this strategy to a formula containing a static target (e.g. $Cancer \sim X + Y + Z$) yields many occlusions, since the CFS algorithm creates the same feature space for every slice.
%%
For such formulas, the 3D view encodes every data element as absolute difference between its $R^2$ features and the global mean along the z-axis.
%%
This highlights slices with unusually low or high results (Fig.~\ref{fig:Application}).
%%
%% TODO: Add reference for disadvantages of 3D information visualizations!
\paragraph{Tackling the disadvantages of 3D information visualization.}
3D information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
%%
We aim to minimize these problems as much as possible.
%%
Since the $R^2$ values are mapped on data point opacity, large values are highlighted in the prism, guiding the focus to the respective slices.
%%
It also creates a very sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
%%
Also, the preceding correlation-based feature selection reduces the information space significantly, leading to sparse cubes.
%%
Overlapping is still an issue, but this way greatly reduced in its affect to the readability of the visualization.
%%

@@

The purpose of this view is the detailed assessment of the underlying regression models.
%%
% TODO: p values kurz nennen? woher kommen die confidence intervals?
By hovering over a data entry in the plot, a tooltip model displays detailed information about a model's coefficients, associated $p$ values and confidence intervals.
%%

@@

\section{Implementation} \label{implementation}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.5in]{figures/implementation}
 \caption{
 The front-end (left) is realized with \emph{HTML5}/\emph{CSS3}/\emph{Javascript} and different Javascript libraries, such as \emph{Angular.js}, \emph{Three.js} and D3.js.
 %%
 The web-server back-end (right) is written in \emph{Node.js}, hosting is carried out using \emph{Heroku}.
 %%
 \emph{R} and \emph{OpenCPU} constitute the statistical back-end (top) to compute the \emph{Regression Cubes}.
 %%
 Additional statistical back-ends can be attached to the system to decrease the computation time.
% The number of attached machines to the system 
%    the  the  Describe the Bootstrap/Angular/D3 Frontend, Node backend, R/OCPU Backend.}
 }
  \label{fig:Implementation}
\end{figure}
%%
% Using web technologies for prototyping allows for:
% \begin{itemize}
% 	\item easy exchange of software, without the need of installing anything other than a modern web browser,
% 	\item fast iterative adaption of method changes, without the need of installing updates,
% 	\item outsourcing computation-heavy tasks to server-clusters, which do not have to be in local proximity.
% \end{itemize}
% %%
%%
We rely on web-based technologies for our prototype.
%% TODO: nochmal Hinweis auf High dimensionality of data?!
The ongoing transition of open-science software into the web spawned numerous projects, making state-of-the-art algorithms available in this domain.
%%
\paragraph{Front-End.}
%%
The front-end is created using \texttt{HTML5}, \texttt{CSS3} and \texttt{Javascript}.
%%
\texttt{Angular.js}\footnote{Open Source; Maintained by Google, \href{https://www.angularjs.org/}{\texttt{angularjs.org}}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
%%
\texttt{Angular.js} is forcing developers to write modularized code, which makes the components easier expandable while keeping the code maintainable by including unit tests.
%%
% UI muss mal erklärt ewrden, tauchte schon weiter vorn im Paper auf
The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of UI elements with proper stylings.
%%
The 2D mosaic plot is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is popular in information visualization using vector graphics \cite{D3}.
%%
It provides fast and easy methods for binding data to graphical elements.
%%
The 3D plot is created using the WebGL-based \texttt{Threejs}\footnote{Open Source; Originally developed by R. Cabello, \href{http://threejs.org}{\texttt{threejs.org}}} library.
%%
We experimented with different ways for achieving the cube representation, including volume rendering, cube primitives for each data point and shader-based solutions.
%%
Open source volume rendering methods are available but do not satisfy our requirements.
%%
Creating a cube primitive for each data point resulted in non-interactive frame-rates for data sets larger than 30 features (creating 30$^3$ cube primitives).
%%
Therefore, we decided to use a shader-based solution by rendering the cube as a sprite-based particle system, allowing to customize color and opacity of every data point.
%%
It also is the fastest solution that we tested.

@@

%% TODO Ergebnisse nachfragen: das ist doch super. Hier noch ein Satz ob die Radiologen das gut finden. ist das nicht richtig geil? Mehr Infos!!!

>>>>

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{175}

@@

%% Paper title.

@@

While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
\begin{enumerate}
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
%%
The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
Overview visualizations of feature relationships as presented by Angelelli et al. \cite{Angelelli} are often focused on relationships between the visualized features.
%%
Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
These target features often indicate the presence of the investigated disease.
%%
As described in Section \ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing these relationships.
%%
% unified -> singular [Steffi?]
A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
\subsection{Cube Description Using Regression Formula Notation}
%%
Expert knowledge is integrated into regression analysis by the regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%
%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
%%

@@

\subsection{Abstracting Regression Results Using $R^2$}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.5in]{figures/cube}
 \caption{
 (left) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The $R^2$ values extracted from the regression formulas depict the quality-of-fit and are mapped to color saturation (a saturated color shows a strong correlation).
 %%
 (right) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The $R^2$ value of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ value is the metric allowing for this kind of assessment.
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between [0,1], it allows for comparison between regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%

@@

\section{System Design} \label{sec:SystemDesign}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.5in]{figures/system}
 \caption{
 %%
 Breast fat data set loaded into the \emph{Regression Cube} prototype.
 %%
 (a) Using the formula input, the user specifies the dependent feature as well as its calculation rules.
 %%
 (b) 3D cube visualization, showing values above the cube matrix diagonal as overview.
 %%
 The values of the currently selected slice are mirrored and represented as orange data points on the slicing plane.
 %%
 (c) 2D mosaic plot visualization of the selected slice for feature \emph{Pain/Discomfort}.
 %%††
 }
  \label{fig:System}
\end{figure}
%% scientific english: kurz und langweilig ;)
We designed our system to be freely accessible and easy to use.
With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
%%
The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses, maybe with other (statistical) tools.
%%
Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.
%%
\\\\
%%
\noindent Design choices and spaces are restricted by the underlying technologies and devices.
%%
Resting method and data exchange on web-based technologies offers many advantages.
%%
There is no set-up time involved and domain experts can use the methods from any computer connected with the web.
%%
By design, web technology is based on a client-server architecture, making it easy to outsource computationally heavy tasks on server clusters and transferring results to the client device.
%%
The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.
%%
Therefore, we have to adapt our design respecting the affordances of web pages.
%%
\subsection{System Paradigm and Components}
%%
The \emph{Regression Cube} design focuses on a clean interface, reducing the amount of user-interface elements as much as possible.
%%
This allows for a fast learning of the individual system parts.
%%
%More importantly, it allows the user to focus all mental resources on the analysis tasks, rather than wrangling with configuring the system to serve the current task or hypothesis.
%%
Our prototype consists of three components:
\begin{itemize}
	%%
	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
	%%
	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube [2.SI].
	%%
	\item The \emph{formula editor} allows to adjust the formula w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
	%%
	It also allows to select a reference formula for creating a difference cube [3.ZF].
	%%
\end{itemize}
%%\com{Focus on simplification and ease of use, reducing the number of UI elements as much as possible (Statistical Visualizations often tend to put all information and settings at once to the user, which is overwhelming for the user).}
\paragraph{File Upload and Classification [1.AF].}
%%
Popular analytics tools, such as WEKA \cite{WEKA}, owe their popularity to their support of open file types.
%%
To allow other users even outside the epidemiological application domain to access to our tool, we use standard ASCII-based CSV files.
%%
The first line in a CSV file represents all features (columns) of the data set.
%%
Each line after that represents one subject (row) and its feature manifestations.

@@

\textbf{Data Security.}
%%
Security issues are raised by uploading data into an online service such as our prototype.
%%
The use of epidemiological data is preceded by a detailed description of the analysis purpose and has to be approved by ethics committees.
%%
Preventive steps have to be taken to restrict access to unauthorized subjects.
%%
%We decided against a user account system because it reduces the ease of use largely without having a direct advantage for the user.
%%
We apply a SHA-256 hashing on the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
%%
Data sets are deleted from the server after closing a session.

@@

\paragraph{3D Prism as Data Mini-Map.}
%%
The 3D cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
%%
It serves as a function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
%%
%It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
%%
The displayed prism shows values above the matrix diagonal.
%%
For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes absolute $R^2$ values (Fig.~\ref{fig:System} b).
%%
Applying this strategy to a formula containing a static target (e.g. $Cancer \sim X + Y + Z$) yields many occlusions, since the CFS algorithm creates the same feature space for every slice.
%%
For such formulas, the 3D view encodes every data element as absolute difference between its $R^2$ features and the global mean along the z-axis.
%%
This highlights slices with unusually low or high results (Fig.~\ref{fig:Application}).
%%
\paragraph{Tackling the disadvantages of 3D information visualization.}
%%
3D information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
%%
We aim to minimize these problems as much as possible.
%%
Since the $R^2$ values are mapped on data point opacity, large values are highlighted in the prism, guiding the focus to the respective slices.
%%
It also creates a very sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
%%
Also, the preceding correlation-based feature selection reduces the information space significantly, leading to sparse cubes.
%%
Overlapping is still an issue, but this way greatly reduced in its affect to the readability of the visualization.
%%

@@

The purpose of this view is the detailed assessment of the underlying regression models.
%%
By hovering over a data entry in the plot, a tooltip model displays detailed information about a model's coefficients, associated $p$ values and confidence intervals.
%%

@@

\section{Implementation} \label{implementation}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.5in]{figures/implementation}
 \caption{
 The front-end (left) is realized with \emph{HTML5}/\emph{CSS3}/\emph{Javascript} and different Javascript libraries, such as \emph{Angular.js}, \emph{Three.js} and D3.js.
 %%
 The web-server back-end (right) is written in \emph{Node.js}, hosting is carried out using \emph{Heroku}.
 %%
 \emph{R} and \emph{OpenCPU} constitute the statistical back-end (top) to compute the \emph{Regression Cubes}.
 %%
 Additional statistical back-ends can be attached to the system to decrease the computation time.
% The number of attached machines to the system 
%    the  the  Describe the Bootstrap/Angular/D3 Frontend, Node backend, R/OCPU Backend.}
 }
  \label{fig:Implementation}
\end{figure}
%%
% Using web technologies for prototyping allows for:
% \begin{itemize}
% 	\item easy exchange of software, without the need of installing anything other than a modern web browser,
% 	\item fast iterative adaption of method changes, without the need of installing updates,
% 	\item outsourcing computation-heavy tasks to server-clusters, which do not have to be in local proximity.
% \end{itemize}
% %%
%%
We rely on web-based technologies for our prototype.
%%
The ongoing transition of open-science software into the web spawned numerous projects, making state-of-the-art algorithms available in this domain.
%%
\paragraph{Front-End.}
%%
The front-end is created using \texttt{HTML5}, \texttt{CSS3} and \texttt{Javascript}.
%%
\texttt{Angular.js}\footnote{Open Source; Maintained by Google, \href{https://www.angularjs.org/}{\texttt{angularjs.org}}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
%%
\texttt{Angular.js} is forcing developers to write modularized code, which makes the components easier expandable while keeping the code maintainable by including unit tests.
%%
% UI muss mal erklärt ewrden, tauchte schon weiter vorn im Paper auf
The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of UI elements with proper stylings.
%%
The 2D mosaic plot is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is popular in information visualization using vector graphics \cite{D3}.
%%
It provides fast and easy methods for binding data to graphical elements.
%%
The 3D plot is created using the WebGL-based \texttt{Threejs}\footnote{Open Source; Originally developed by R. Cabello, \href{http://threejs.org}{\texttt{threejs.org}}} library.
%%
We experimented with different ways for achieving the cube representation, including volume rendering, cube primitives for each data point and shader-based solutions.
%%
Open source volume rendering methods are available but do not satisfy our requirements.
%%
Creating a cube primitive for each data point resulted in non-interactive frame-rates for data sets larger than 30 features (creating 30$^3$ cube primitives).
%%
Therefore, we decided to use a shader-based solution by rendering the cube as a sprite-based particle system, allowing to customize color and opacity of every data point.
%%
It also is the fastest solution that we tested.

@@

%% TODO Feedback of Henry!
